\documentclass[a4paper,12pt]{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{array}
\usepackage{longtable}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{setspace}
\usepackage{color}
\pagestyle{fancy}
\fancyhf{}
\rhead{ET23BTIT133}
\renewcommand{\headrulewidth}{0.4pt}

\begin{document}

\begin{center}
    {\LARGE \textbf{Advanced Web Technologies}} \\[6pt]
    {\large Practical Work and Lab Assignments} \\[16pt]
    \begin{tabular}{ll}
        \textbf{Name:} Sree Saicharan Vadapalli & \hspace{5em} \textbf{Subject Code:} BTIT13601 \\
        \textbf{Branch:} Information Technology & \hspace{5em} \textbf{Year:} B. Tech III (Sem VI) \\
        \textbf{Enrollment No:} ET23BTIT133 & \\
    \end{tabular}
\end{center}

\vspace{5pt}

\begin{center}
    {\large \textbf{INDEX}}
\end{center}

\vspace{6pt}

\begin{longtable}{|>{\centering\arraybackslash}p{1cm}|>{\centering\arraybackslash}p{2cm}|p{8.5cm}|>{\centering\arraybackslash}p{1.2cm}|>{\centering\arraybackslash}p{2cm}|}
\hline
\textbf{Sr. No.} & \textbf{Date} & \centering\textbf{Experiments} & \textbf{Page No.} & \textbf{Signature} \\
\hline
\endfirsthead
\hline
\textbf{Sr. No.} & \textbf{Date} & \centering\textbf{Experiments} & \textbf{Page No.} & \textbf{Signature} \\
\hline
\endhead

1 & & Create a registration form and apply CSS and perform at least 5 JavaScript validations. & & \\[40pt]
\hline

2 & & Create a simple guess for the number type game. It should choose a random number between 1 and 100, then challenge the player to guess the number in 10 turns. After each turn the player should be told if they are right or wrong, and if they are wrong, whether the guess was too low or too high. It should also tell the player what numbers they previously guessed. The game will end once the player guesses correctly, or once they run out of turns. When the game ends, the player should be given an option to start playing again. & & \\[10pt]
\hline

3 & & Create a Hello World Web Page using React JS. & & \\[40pt]
\hline

4 & & Implement a react app to increment and decrement the Count with useState hook. & & \\[30pt]
\hline

5 & & Implement a react app to calculate the area of a circle. & & \\[40pt]
\hline

6 & & Create a Sign-Up Form in React JS. The Sign-up form should ask for Name, Mobile, Email, Address, Date of Birth, Gender, Username, Password and Confirm Password. The Form should have two buttons, one for reset and one for submitting. The Submit Button should be enabled only when all the input values are validated. & & \\[10pt]
\hline

7 & & Implement a react app to pass the message ``Hello from Parent'' from parent to child component using Props. & & \\[30pt]
\hline

8 & & Develop a simple application using React.js where a user can Add/Delete notes. Each note time stamped as well. & & \\[30pt]
\hline

9 & & Develop an application to display ``Hello, SCET!'' when the button is clicked. Use React Context (create and export the context in a separate file.) & & \\[30pt]
\hline

10 & & Create a Hello World Nodejs Application. & & \\[40pt]
\hline

11 & & Create an Http Server which will respond with the message ``Welcome to the World of Nodejs'' to the client. & & \\[30pt]
\hline

12 & & Create an HTTP Server which will respond with current date and time to the client. & & \\[30pt]
\hline

13 & & Demonstrate the working of cookies and sessions in nodejs. & & \\[40pt]
\hline

14 & & Create a Signup Web Application using React JS, nodejs and express framework. The Sign-up form should ask for Name, Mobile, Email, Address, Date of Birth, Gender, Username, Password and Confirm Password. The Form should have two buttons, one for reset and one for submitting. The Submit Button should be enabled only when all the input values are validated. Upon Clicking the Submit Button Your webpage should traverse to a route ``/process\_request'' which is defined using NodeJS and express framework and which will display the contents back to the client and insert all the data to the database made in mongoDB. & & \\[6pt]
\hline

15 & & Create a Single Page Application using MERN Stack. & & \\[40pt]
\hline

\end{longtable}

\end{document}
